Search for Spatially Extended Fermi-LAT Sources Using Two Years of Flight Data

### 2011-06-01 extended_catalog/paper/extended_catalog.tex
More to the draft

\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\newcommand{\tsext}{\ensuremath{\text{TS}_\text{ext}}\xspace}
\newcommand{\ts}{\text{TS}\xspace}
\newcommand{\pointlike}{{\em pointlike}\xspace}
\newcommand{\gtlike}{{\em gtlike}\xspace}

@@

\begin{abstract}
We present a new method for analyzing spatially extended sources with
the Fermi Large Area Telescope (LAT), the primary science instrument
on the {\em Fermi Gamma-ray Space Telescope (Fermi)}.  We provide a
series of Monte Carlo studies that were done to validate this tool.
We then calculate the Fermi LAT's sensitivity to spatially exnteded
sources.
We then apply this tool to test all sources in the 2FGL Two Year Catalog
for extension.\cite{2FGL} We then report on the discovery of XXX new
extended sources.
\end{abstract}

@@

Nevertheless, morphological studies of sources in the $\gev$ energy range using
the Fermi LAT is challenging primarily because of the significantly energy
dependent PSF and because of systematic errors in the
galactic diffuse emission. The Fermi PSF varies from over $XXXXX^\circ$
at $100\mev$ to $XXXXX^\circ$ at $100\gev$ and so the higher energy photons are

@@

Performing morphological studies of sources in the $\gev$ energy range using
the Fermi LAT is challenging primarily because of the significantly energy
dependent PSF and because of systematic errors in the
galactic diffuse emission. The Fermi PSF varies from over $XXXXX^\circ$
at $100\mev$ to $XXXXX^\circ$ at $100\gev$ and so the higher energy photons are
significantly more important for studying the shape of sources. One could
ideally look just at the highest energy photons, but the trade off is that
for typical sources many more source events are expected at low energy.
For this reason, it is especially important to use as much of the data
as possible.

@@

@@

@@

To calculate the statistical significance of the extension of a source,
one can define a parameter $\tsext$ as twice the increase
in log likelihood between fitting a source with a
spatially extended hypothesis and a point hypothesis. 
So $\tsext=2\times(LL_\text{ext}-LL_\text{ps})$.

@@

@@

@@

Fermi LAT data was simulated using the publically avaliabable Science
Tools package {\em gtobssim}\cite{GTOBSSIM_CITATION}. The sources were
simulated with a powerlaw spectral model with 6 Were given 6 $>100\mev$
fluxes ranging from $3\times 10^{-9} \ph/\cm^2/s$ to 
$10^{-6} \ph/\cm^2/s$.
The sources were then simulated with spectral index ($\gamma$) values of
1.5, 2, 2.5, and 3.  These values were picked to probe a representative
range measurable values.  For particular spectral indices, only the
fluxes which would create a statistically significant source detection
($\ts>25$) where selected, since we are only interested in
testing for extension sources with a significant extent. The sources were
simulated on top of an isotropic powerlaw background with a Sreekumar-like
spectrum ($>100\mev$ flux of 1.5e-5 and spectral index 2.1).\cite{Sreekumar
et al. ApJ 494 pag 523 1998}

@@

The results of the monte carlo simulation are shown in
figure~\ref{ts_ext_mc}.  The cumulative density function (CDF) is plotted
against $\tsext$ For reference, the CDF for $\chi^2_1/2$,
suggested by Wilk's Theorem, is also plotted for comparison. Overall,
there is good agreement between Wilk's theorem, although the agreement
is not perfect.  Possible reasons for departure from Wilk's theorem may
include \pointlike ignoring energy dispersion, which would generally change
the shape as a function of energy, numerical innacuracies in evaluating
the convolution algorithm, or possibly problems with the fitting algorithm
getting stuck in false minimum. It is important to emphasize that if
our emperical curve lies to the left of the curve we assume to estimate
significance, we would underestimate the statistical significance of
our result. Since this is the case for most of our simluations, we are
reasonable confident that $\sqrt{\tsext}$ can be safely
used as a measure of the statistical significance of a source's extension.

@@

Our first result is the extension sensitivity for sources of varying
spectral indices. The result is shown in figure~\ref{index_sensitivity}.
We computed the sensitivty for sources with a spectral index of 1.5,
2, 2.5, and 3 and for sources of 20 extensions varying from $0.1^\circ$
to $2.0^\circ$. The $y$ axis is the $100\mev$ to $100\gev$ flux of the
source and the $x$ axis is the monte carlo extension. The curves then
represent the division in this extension and flux space where one gets
on average $\tsext=25$.
The solid lines of the plot are the sensitivity computed when using
with all events from $100\mev$ to $100\gev$. The dashed lines is the
same sensitivity but only using events from from $1\gev$ to $100\gev$.

This results shows that our extension sensitivity is
significantly better for harder sources where there are more high
energy events. Furthermore, our sensitivity is best for sources with an
extension around $0.5^\circ$.  our sensitivity gets significantly worse
for very small sources and gets slightly worse for very large sources.

@@

For each source, we used a circular $10^\circ$ region of interest centered
on our source. We performed a fully binned analysis.  The spatial bin
size was scaled with energy as was described above and we uniformly
binned in log energy, using 4 binned per logarithmic decade of of energy.
We included all background catalog sources within $15^\circ$ of the source
of interest and refit the spectral parameters of sources within $2^\circ$
of our source.

@@

First, we took the spatial and spectral model found using \pointlike
and performed a crosscheck by refitting the spectral parameters using 
\gtlike.
We used a square ROI $10^\circ$ in diamater, the same 4 degrees
per decade, and $0.125^\circ$ energy binning.
The \gtlike analysis provided a second estimate of the spectral
paramters of the extended source as well as a second calculation of
$\ts_\text{disk}$, $\ts_\text{point}$, and 
$\tsext$. We only looked for extended sources which 
both \pointlike and \gtlike considered significantly extended.

@@

Following the Fermi-LAT one year catalog, the Fermi collaboration
published a list of the Fermi catalog sources that had a high probability
association with AGN. The first year catalog had 709 AGN associated with
671 distinct 1FGL sources at high latitude ($|b|>10^\circ$).  For our
study, we are looking to assess systematics with extension fitting so
it is important to avoid any possible issues with AGN associations.
For that reason, we select only the 599 AGN which make it into the
``clean'' AGN sample. An AGN association is only considered ``clean''
if it has a high probability of association $P\ge 80\%$, if it is the
only AGN associated with the 1FGL source, and if there are no flags
on the source which shed doubt on the source\cite{1FGL}. These last two
conditions are especially important for our analysis. One might expect
source confusion between two AGN look like a spatially extended source.
Testing only unflagged sources is also important because it removes
any catalog sources coincident with structures in the diffuse emission.
It is possible that these flagged catalog sources which are really poorly
modeled diffuse emission might also look extended.

@@

Using the same extension fitting procedure as described above, we test
these 515 AGN for extension. A histogram of the fit \tsext values is
shown in figure~\ref{agn_ts_ext}. Overlaid on the plot is again
a $\chi^2/2$ distribution suggested by Wilk's Theorem.
This plot shows a worse agreement with the theoretical distribution
then was found by our monte carlo study in figure~\ref{ts_ext_mc}.

@@

The third possible effect causing this discrepancy is an
imperfect modeling of the PSF of the LAT.  Before launch,
the LAT PSF was determined using a Monte Carlo simulation of
the detector.  After launch, a discrepancy was found using flight
data above a few \gev in the PSF measured using very bright AGN.
This result was presented in conference and a publication is in preparation.
Subsequently, the PSF was fit empirically to these bright AGN and
this empirical parameterization is used in the Pass 7 Version 6 IRFs.\cite{
https://confluence.slac.stanford.edu/download/attachments/102860834/FermiSymp2011_CAPSF_v5_ROTH.pptx
}. But our current parameterization is not perfect. There are 
especially lingering issues because there is not good enough statistics
at high energy to validate the PSF as a function of the incident $\theta$
angle of a photon. It is likely that these issues could cause this 
systematic in the fit value of \tsext for point sources. We should
furtheremore note that since the PSF was

with only 2 AGN having $\tsext>10$ and the maximum having $\tsext=13.7$

@@

\section{Interpretation}

@@

\begin{table}
  \begin{centering}
\begin{tabular}{ | r | r | r | r | }
\hline
$\gamma$ & flux ($\ph/\cm^2/\s$) & $N_\text{sims}$ & $<\ts>$ \\
\hline
  1.5 &          $10^{-6}$ &           11405 &  90861 \\
      &  $3\times 10^{-7}$ &           11490 &  21922 \\
      &          $10^{-7}$ &           11101 &   5765 \\
      &  $3\times 10^{-8}$ &           10269 &   1262 \\
      &          $10^{-8}$ &            9941 &    306 \\
      &  $3\times 10^{-9}$ &           10427 &     61 \\
\hline
    2 &          $10^{-6}$ &           11632 &  22025 \\
      &  $3\times 10^{-7}$ &           11622 &   4886 \\
      &          $10^{-7}$ &           11115 &   1159 \\
      &  $3\times 10^{-8}$ &           10503 &    222 \\
      &          $10^{-8}$ &            9893 &     50 \\
\hline
  2.5 &          $10^{-6}$ &           11522 &   4722 \\
      &  $3\times 10^{-7}$ &           11445 &    948 \\
      &          $10^{-7}$ &           10966 &    202 \\
      &  $3\times 10^{-8}$ &            8227 &     41 \\
\hline                                                
    3 &          $10^{-6}$ &           11264 &    949 \\
      &  $3\times 10^{-7}$ &           10986 &    172 \\
      &          $10^{-7}$ &            7855 &     40 \\
\hline
\end{tabular}
\caption{
This table presents a list of the spectral models (flux and spectral index)
used in the monte carlo study of 
\pointlike described in section~\ref{monte_carlo_validation}.
For each spectral model, the number of statistically indepdent simulations
and the average test statistic found for that given spectral model
($<\ts>$) is tabulated. 
The important feature to see in this table is that the monte carlo simulation
probes most of the sources parameters that could be detected by the LAT.
The simulation probles a wide range of spectral indices (from index 1.5 to 3) and
a wide range of the fluxes (varying
from a close to $5\sigma$ source detection to a $>10\sigma$ source detection).
Only simluations which produced statistically
significant detections ($\ts>25$) were kept. 
What is quoted is the $>100\mev$ integral flux in measured in units of $\ph/\cm^2/\s$.
}
\label{ts_ext_num_sims}
  \end{centering}
\end{table}

@@

\clearpage
\begin{figure}
  \begin{center}
    \includegraphics{ic443_plots/ic443_ts_ext_vs_energy.pdf}
    % taken from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/ts_ext_vs_energy/v1/ts_ext_vs_energy.py
    \caption{\tsext vs log(energy)}
    \label{counts_slice}
  \end{center}
\end{figure}

@@

\clearpage
\begin{figure}
  \begin{center}
    \includegraphics{ic443_plots/ic443_smoothed_counts.pdf}
    % taken from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/smoothed_counts_ic443/v1
    \caption{Smoothed counts}
    \label{smoothed_counts}
  \end{center}
\end{figure}

\clearpage
\begin{sidewaystable}
    \begin{centering}
      \begin{tabular}{rrrrrrrrrrr}
        Name           &    GLON &    GLAT & $\sigma(^\circ)$ &     \ts & $\tsext$ & Major & Minor &   Ang & Flux ($10^{-9}$) &         Index \\
        \hline
        \multicolumn{11}{l}{$E > 1\gev$} \\
        \hline
        SMC            & 302.87  &  -44.65 &    $1.59\pm0.15$ &    93.1 &                   51.2 & 0.173 & 0.129 &  41.9 &   $3.0\pm0.4$    & $2.42\pm0.16$ \\
        LMC            & 278.58  &  -32.82 &    $2.26\pm0.09$ &   384.2 &                304.374 & 0.096 & 0.071 & -35.5 &  $12.9\pm0.7$    & $2.42\pm0.08$ \\
        IC443          & 189.05  &  3.04   &    $0.35\pm0.01$ & 10765.3 &                535.282 & 0.006 & 0.006 &  84.0 &  $65.2\pm1.2$    & $2.23\pm0.02$ \\
        W28            & 6.51    &  -0.29  &    $0.39\pm0.02$ &  1231.3 &                   92.1 & 0.014 & 0.013 &  30.4 &  $55.9\pm1.8$    & $2.65\pm0.03$ \\
        W30            & 8.61    &  -0.20  &    $0.36\pm0.04$ &   458.1 &                  66.0  & 0.020 & 0.018 &  14.1 &  $30.0\pm1.8$    & $2.58\pm0.06$ \\
        W44            & 34.68   &  -0.42  &    $0.34\pm0.02$ &  1387.6 &                  111.3 & 0.009 & 0.009 & -39.4 &  $74.7\pm1.0$    & $2.67\pm0.01$ \\
        W51C           & 49.12   &  -0.44  &    $0.28\pm0.02$ &  2174.7 &                   87.6 & 0.010 & 0.010 &  59.4 &  $41.6\pm1.3$    & $2.38\pm0.04$ \\
        \hline
        \multicolumn{11}{l}{$E > 10\gev$} \\
        \hline
        HESS J1825-137 &   17.58 &   -0.46 &    $0.64\pm0.08$ &    85.6 &                   62.5 & 0.050 & 0.045 & 33.6  &     $19.4\pm1.4$ & $2.27\pm0.07$ \\
        MSH 15-52      &  320.39 &   -1.22 &    $0.20\pm0.06$ &    73.7 &                    9.5 & 0.033 & 0.032 & 3.9   &      $4.7\pm0.7$ & $1.88\pm0.12$ \\
      \label{known_extended_sources}
      \end{tabular}
      \caption{The fit results for the known extended sources that
      were found by our analysis. The top results were found using
      an energy range between $1\gev$ and $100\gev$.
      The quoted flux is in units of $10^{-9}\ph/\cm^2/\s$ and 
      is the flux in the fit energy range.
      The lower analysis was done using events between
      $10\gev$ and $100\gev$ and the quoted flux
      values are in the fit energy range. All sources were
      fit using a disk morphology. The localiztaion and extension
      were found using \pointlike while the TS and spectral
      values were calculated using \gtlike.
      The TS value was computed using 
      the assumed disk morphology. The Centaurus A Lobes,
      the Cygnus Loop, and Vela X were not found to be extended by
      this too. The reasons for this are discussed in~\ref{validate_known}.
      }
    \end{centering}
\end{sidewaystable}

>>>>

% some units
\newcommand{\mev}{\text{MeV}\xspace}
\newcommand{\gev}{\text{GeV}\xspace}
\newcommand{\sr}{\text{sr}\xspace}
\newcommand{\s}{\text{s}\xspace}
\newcommand{\ph}{\text{ph}\xspace}
\newcommand{\cm}{\text{cm}\xspace}
\newcommand{\tsext}{\ensuremath{\text{TS}_\text{ext}}\xspace}
\newcommand{\tsinc}{\ensuremath{\text{TS}_\text{inc}}\xspace}
\newcommand{\ts}{\text{TS}\xspace}

\renewcommand{\deg}{\ensuremath{^\circ}\xspace}

% the program names
\newcommand{\pointlike}{{\em pointlike}\xspace}
\newcommand{\gtlike}{{\em gtlike}\xspace}
\newcommand{\gtobssim}{{\em gtobssim}\xspace}

@@

\begin{abstract}
We present a new method for analyzing spatially extended sources with
the Fermi Large Area Telescope (LAT), the primary science instrument
on the {\em Fermi Gamma-ray Space Telescope (Fermi)}.  We provide a
series of Monte Carlo studies that were done to validate this tool.
We then calculate the Fermi LAT's sensitivity to spatially exnteded
sources.
We then apply this tool to test all sources in the 2FGL Two Year Catalog
for extension.\cite{2FGL} We then report on the discovery of XXXXXXXXXXXXXX new
extended sources.
\end{abstract}

@@

Nevertheless, morphological studies of sources in the $\gev$ energy range using
the Fermi LAT is challenging primarily because of the significantly energy
dependent PSF and because of systematic errors in the
galactic diffuse emission. The Fermi PSF varies from over $XXXXXXXXXXXXXXX\deg$
at $100\mev$ to $XXXXXXXXXXXXXXXX\deg$ at $100\gev$ and so the higher energy photons are

@@

Performing morphological studies of sources in the $\gev$ energy range using
the Fermi LAT is challenging primarily because of the significantly energy
dependent PSF and because of systematic errors in the
galactic diffuse emission. The Fermi PSF varies from over $XXXXXXXXXXXXXXXX\deg$
at $100\mev$ to $XXXXXXXXXXXXXXXX\deg$ at $100\gev$ and so the higher energy photons are
significantly more important for studying the shape of sources. One could
ideally look just at the highest energy photons, but the trade off is that
for typical sources many more source events are expected at low energy.
For this reason, it is especially important to use as much of the data
as possible.

@@

\subsection{\pointlike}

@@

% XXXXXXXXXXXX describe how \pointlike calculates extension errors and how localization errors are calculated.

% XXXXXXXXXXXX note about how \pointlike is written in python = more flexible

\subsection{Extension Fitting}

@@

To calculate the statistical significance of the extension of a source,
one can define a parameter $\tsext$ as twice the increase
in log likelihood between fitting a source with a
spatially extended hypothesis and a point hypothesis. 
So 
\begin{equation}
  \tsext=2\times(LL_\text{ext}-LL_\text{ps}).
\end{equation}

@@

% XXXXXXXXXXXXXXXXXXX NOTE ABOUT gtlike crosscheck

\subsection{Dual Localization}

There is a degeneracy in the signal expected from spatially extended
sources and multiple point sources.  To assess the possibility of source
confusion, we developed a function in pointlike for simulatenously fitting
two point sources so that the likelihood fitting two point sources
could be compared to the likelihood fitting one extended source.

The algorithm works by first dividing the selected point source into
two separate sources. Then, the position of the two sources 
are fit simultanously. To avoid projection effects, the fit is
done in a coordiante system rotated to the celestial equator.
In this rotated coordiante system, space can be well approximated
as being flat and we are interested in fitting the 
rotated coordiante $(x_1',y_1')$ and $(x_2',y_2')$
We decided not to directly fit the positions of the two sources 
but instead the fit sum and difference of the rotated coordiantes
$S_x=(x_2'+x_1')/2$, $S_x=(y_2'+y_1')/2$, $D_x=(x_2'-x_1')/2$, and $D_y=(y_2'-y_1')/2$.
We believe that fitting the average of position and the difference
independently would cause less degeneracy in the fit and help the fitter
find the best position.

For a particular value of $S_x$, $S_y$, $D_x$, and $D_y$, the Poisson
likelihood is computed by converting the sum and difference coordaintes
into the rotated coordiantes, unrotating the coordiantes, moving the
two sources in the \pointlike to the new positions, and performing a new
spectral analysis at the modified position.  The function then maximizes
The Poission likelihood $L(S_x,S_y,D_x,D_y)$ as a function of the sum
and difference of the rotated coordaintes using the Minuit to fit the
sum and differenc coordiantes\cite{Minuit reference}.

After the minimization converges, we can define the 
parameter $\tsext$ as twice the 
increase
in log likelihood between fitting the region with
two point sources and fitting the region with one point source
\begin{equation}
  \tsext=2\times(LL_\text{ext}-LL_\text{ps}).
\end{equation}
Generally, \tsinc can also

% XXXXXXXXXXXXXXX \tsinc vs \tsext NOT A CALIBRATED TEST
% but still suggestive in the cases \tsinc \ll \tsext and
% \tsinc \gg \tsext.

Because of \pointlike's computational efficiency, this algorithm
runs in a reasonable amount of time.

After
% XXXXXXXXXXXXXXXXXXX NOTE ABOUT gtlike crosscheck

\subsection{Comparing Source Sizes}

Description of comparing r68, sigma for disk, gaussian, etc.
why the plots don't look correct.

@@

% XXXXXXXXXXXXXXXX better description of how TSext is always > 0
% but wilk's thorem says that it won't be must better if the source
% is a point source

@@

Fermi LAT data was simulated using the publically avaliabable Science
Tools package \gtobssim\cite{GTOBSSIM_CITATION}. The sources were
simulated with a powerlaw spectral model with 6 Were given 6 $>100\mev$
fluxes ranging from $3\times 10^{-9} \ph/\cm^2/s$ to 
$10^{-6} \ph/\cm^2/s$.
The sources were then simulated with spectral index ($\gamma$) values of
1.5, 2, 2.5, and 3.  These values were picked to probe a representative
range measurable values.  For particular spectral indices, only the
fluxes which would create a statistically significant source detection
($\ts>25$) where selected, since we are only interested in
testing for extension sources with a significant extent. The sources were
simulated on top of an isotropic powerlaw background with a Sreekumar-like
spectrum ($>100\mev$ flux of 1.5e-5 and spectral index 2.1).\cite{Sreekumar
et al. ApJ 494 pag 523 1998}

@@

The results of the monte carlo simulation are shown in
figure~\ref{ts_ext_mc}.  The cumulative density function (CDF) is plotted
against $\tsext$ For reference, the CDF for $\chi^2_1/2$,
suggested by Wilk's Theorem, is also plotted for comparison. Overall,
there is good agreement between Wilk's theorem, although the agreement is
not perfect.  It should be noted that the discrepancy seems to be worst
for bright sources which seems to imply that numerical errors in the
way \pointlike models the PSF and performs the convolution become more
apparent for sufficiently bright sources.  Other possible reasons for
departure from Wilk's theorem might include \pointlike ignoring energy
dispersion. Energy dispersion will change the shape as a function of
energy. \pointlike may also run into possibly problems with the fitting
algorithm getting stuck in false minimum. It is important to emphasize
that if
our emperical curve lies to the left of the curve we assume to estimate
significance, we would underestimate the statistical significance of
our result. Since this is the case for most of our simluations, we are
reasonable confident that $\sqrt{\tsext}$ can be safely
used as a measure of the statistical significance of a source's extension.

@@

Our first result is the extension sensitivity for sources of varying
spectral indices. The result is shown in figure~\ref{index_sensitivity}.
We computed the sensitivty for sources with a spectral index of 1.5,
2, 2.5, and 3 and for sources of 20 extensions varying from $0.1\deg$
to $2.0\deg$. The $y$ axis is the $100\mev$ to $100\gev$ flux of the
source and the $x$ axis is the monte carlo extension. The curves then
represent the division in this extension and flux space where one gets
on average $\tsext=25$.
The solid lines of the plot are the sensitivity computed when using
with all events from $100\mev$ to $100\gev$. The dashed lines is the
same sensitivity but only using events from from $1\gev$ to $100\gev$.

This results shows that our extension sensitivity is
significantly better for harder sources where there are more high
energy events. Furthermore, our sensitivity is best for sources with an
extension around $0.5\deg$.  our sensitivity gets significantly worse
for very small sources and gets slightly worse for very large sources.

@@

For each source, we used a circular $10\deg$ region of interest centered
on our source. We performed a fully binned analysis.  The spatial bin
size was scaled with energy as was described above and we uniformly
binned in log energy, using 4 binned per logarithmic decade of of energy.
We included all background catalog sources within $15\deg$ of the source
of interest and refit the spectral parameters of sources within $2\deg$
of our source.

@@

First, we took the spatial and spectral model found using \pointlike
and performed a crosscheck by refitting the spectral parameters using 
\gtlike.
We used a square ROI $10\deg$ in diamater, the same 4 degrees
per decade, and $0.125\deg$ energy binning.
The \gtlike analysis provided a second estimate of the spectral
paramters of the extended source as well as a second calculation of
$\ts_\text{disk}$, $\ts_\text{point}$, and 
$\tsext$. We only looked for extended sources which 
both \pointlike and \gtlike considered significantly extended.

@@

Following the Fermi-LAT one year catalog, the Fermi collaboration
published a list of the Fermi catalog sources that had a high probability
association with AGN. The first year catalog had 709 AGN associated with
671 distinct 1FGL sources at high latitude ($|b|>10\deg$).  For our
study, we are looking to assess systematics with extension fitting so
it is important to avoid any possible issues with AGN associations.
For that reason, we select only the 599 AGN which make it into the
``clean'' AGN sample. An AGN association is only considered ``clean''
if it has a high probability of association $P\ge 80\%$, if it is the
only AGN associated with the 1FGL source, and if there are no flags
on the source which shed doubt on the source\cite{1FGL}. These last two
conditions are especially important for our analysis. One might expect
source confusion between two AGN look like a spatially extended source.
Testing only unflagged sources is also important because it removes
any catalog sources coincident with structures in the diffuse emission.
It is possible that these flagged catalog sources which are really poorly
modeled diffuse emission might also look extended.

@@

Using the same extension fitting procedure as described above, we
test these 515 AGN for extension. A histogram of the fit \tsext values
is shown in figure~\ref{agn_ts_ext}. Overlaid on the plot is again a
$\chi^2/2$ distribution suggested by Wilk's Theorem.  This plot shows
a worse agreement with the theoretical distribution then was found
by our monte carlo study in figure~\ref{ts_ext_mc}.  But overall, the
distribution is good. No AGN get fit with a particularly significant
value of $\tsext$. Almost all of the AGN get fit with $\tsext<1$ and
there are very few AGN fit with large values of \tsext.  There where
only 2 AGN having $\tsext>10$.  The maximum fit $\tsext=13.7$ whereas the
maximum value expected theoeritcally for the $\chi^2/2$ distribution is
8.3. This result therefore shows that there is some systematic on our
analysis but that we understand it and that it will not significantly
cast into doubt our subsequent results.
% 8.3 number gotten from mathematica:
% PDF[ChiSquareDistribution[1], x]
% N[Solve[(1/2)*
%     Integrate[1/(E^(x/2)*(Sqrt[2*Pi]*Sqrt[x])), {x, TS, Infinity}] &=&  
%        1/515, TS]]

@@

Another possible systematic may be due to remaning issues with our
background model. Due to chance coincidence, a catalog source might
lie very close to nearby source. If the two sources were near enough,
they would be unresolvable by the catalog and the catalog would simply
fit them both together.  This could easily lead to an anomolously high
value of \tsext for a paritcular point source. Issues could also arise
from farther away but fainter sources which did not make the catalog's
significance cut, but nonetheless would pull the fitter away from the
point hyoptehsis.  Finally, it is possible that there are lingering
residuals in the diffuse emission in parts of the high latitude sky
which would cause the fitter to prefer a slightly extneded hypotehsis
that could fill in parts of the unmodeled diffuse emission.

The final possible effect causing this discrepancy is an imperfect
modeling of the PSF of the LAT.  Before launch, the LAT PSF
was determined using a Monte Carlo simulation of the detector.
After launch, a discrepancy was found using flight data above a
few \gev in the PSF measured using very bright AGN.  This result
was presented in conference and a publication is in preparation.
Subsequently, the PSF was fit empirically to these bright AGN and this
empirical parameterization is used in the Pass 7 Version 6 IRFs.\cite{
https://confluence.slac.stanford.edu/download/attachments/102860834/FermiSymp2011_CAPSF_v5_ROTH.pptx
}. But our current parameterization is not perfect. There are especially
lingering issues due to lackign the statistics at high energy to
validate the PSF as a function of the incident $\theta$ angle of a
photon. Since the distribution of $\theta$ angles for observing any
sources is different, the PSF for every source should be different.
It is likely that these issues could cause this systematic in the fit
value of \tsext for AGN.

We should furtheremore note that since the PSF was found empirically in
the first place by fitting the shape of brigh AGN, finding that the AGN we
test are not significantly extended is not surprising.  We hope to note
that because of the difficulties in validating the PSF, we are strictly
searching for sources that are significantly extended when compared to
bright AGN measured by the LAT. Trying to look for the extension of these
AGN, or more realistically trying to put an upper limit on thier size,
would require its own dedicated analysis.

@@

\begin{itemize}
  \item Results summarized in table~\ref{new_ext_srcs}.
  \item Add extra stuff to this section, $\tsinc$.
  \item Disribe what had to be done to modify regionregion.
    What sources had to be deleted (or moved).
%  \item XXXXXXXXXXXXXXXXXXXXXXXXXX Say that errors are just statistical (or have a descritpion of errors).

\end{itemize}

\subsection{1FGL J1628.6-2419c}

The two year catalog source P72Y2516 was found at ($l,b)=(353.09\deg, 16.79\deg)$
with a best fit extension of $0.40\pm0.05$ and an extension significance of
$\tsext=26.3$.

In the ophiucus region=most likely a molecular cloud

\subsection{1FGL J1554.0-5345c}
Unidentified source likely diffuse emission

\subsection{1FGL J0823.3-4248}
Coincident with Puppis A SNR

\subsection{1FGL J1613.6-5100c}
Near HESS Source XXXXXXXXXXXXXX

\subsection{1FGL J1632.9-4802c}
Near HESS Source XXXXXXXXXXXXXX

\subsection{1FGL J2020.0+4049}
Gamam Cygni SNR. Upcomming publication.

\subsection{1FGL J1837.5-0659c}
Near HESS source HESS J1837. Upcomming publication???

\subsection{P72Y1287}

Vela Jr. Upcomming publication.

\section{Conclusion}

@@

\begin{table}
  \begin{centering}
\begin{tabular}{ | r | r | r | r | }
\hline
$\gamma$ & flux ($\ph/\cm^2/\s$) & $N_\text{sims}$ & $<\ts>$ \\
\hline
  1.5 &          $10^{-6}$ &           31952 &  92862 \\
      &  $3\times 10^{-7}$ &           31962 &  22169 \\
      &          $10^{-7}$ &           31977 &   5806 \\
      &  $3\times 10^{-8}$ &           31991 &   1270 \\
      &          $10^{-8}$ &           31940 &    301 \\
      &  $3\times 10^{-9}$ &           30324 &     62 \\
\hline
    2 &          $10^{-6}$ &           31872 &  22067 \\
      &  $3\times 10^{-7}$ &           31890 &   4898 \\
      &          $10^{-7}$ &           31858 &   1097 \\
      &  $3\times 10^{-8}$ &           31632 &    236 \\
      &          $10^{-8}$ &           27491 &    103 \\
\hline
  2.5 &          $10^{-6}$ &           31822 &   4706 \\
      &  $3\times 10^{-7}$ &           31822 &    889 \\
      &          $10^{-7}$ &           31169 &    176 \\
      &  $3\times 10^{-8}$ &           21591 &     41 \\
\hline                                                
    3 &          $10^{-6}$ &           31763 &    929 \\
      &  $3\times 10^{-7}$ &           31665 &    161 \\
      &          $10^{-7}$ &           19271 &     40 \\
\hline
\end{tabular}
\caption{
This table presents a list of the spectral models (flux and spectral
index) used in the monte carlo study of \pointlike described in
section~\ref{monte_carlo_validation}.  For each spectral model, the
number of statistically indepdent simulations and the average test
statistic found for that given spectral model ($<\ts>$) is tabulated.
The important feature to see in this table is that the monte carlo
simulation probes most of the sources parameters that could be detected
by the LAT.  The simulation probles a wide range of spectral indices
(from spectral index $\gamma$ of 1.5 to 3) and a wide range of the fluxes
(varying from a close to $5\sigma$ source detection to a $>10\sigma$
source detection).  Only simluations which produced statistically
significant detections ($\ts>25$) were kept.  What is quoted is the
$>100\mev$ integral flux measured in units of $\ph/\cm^2/\s$.
}
\label{ts_ext_num_sims}
  \end{centering}
\end{table}

@@

\clearpage
\begin{figure}
  \begin{center}
    \includegraphics{ic443_plots/ic443_ts_ext_vs_energy.pdf}
    % taken from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/ts_ext_vs_energy/v1/ts_ext_vs_energy.py
    \caption{A plot of \tsext vs log(energy)
    for the SNR IC443, which has a typical spectral model.}
    \label{counts_slice}
  \end{center}
\end{figure}

@@

\clearpage
\begin{figure}
  \begin{center}
    \includegraphics{ic443_plots/ic443_smoothed_counts.pdf}
    % taken from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/smoothed_counts_ic443/v1
    \caption{Smoothed counts. Galactic and isotoropic as well as earth albedo background were subtracted.}
    \label{smoothed_counts}
  \end{center}
\end{figure}

\clearpage
\begin{sidewaystable}
    \begin{centering}
      \begin{tabular}{rrrrrrrrrrr}
        Name           &    GLON &    GLAT & $\sigma(\deg)$ &     \ts & $\tsext$ & Major & Minor &   Ang & Flux ($10^{-9}$) &         Index \\
        \hline
        \multicolumn{11}{l}{$E > 1\gev$} \\
        \hline
        SMC            & 302.87  &  -44.65 &    $1.59\pm0.15$ &    93.1 &                   51.2 & 0.173 & 0.129 &  41.9 &   $3.0\pm0.4$    & $2.42\pm0.16$ \\
        LMC            & 278.58  &  -32.82 &    $2.26\pm0.09$ &   384.2 &                304.374 & 0.096 & 0.071 & -35.5 &  $12.9\pm0.7$    & $2.42\pm0.08$ \\
        IC443          & 189.05  &  3.04   &    $0.35\pm0.01$ & 10765.3 &                535.282 & 0.006 & 0.006 &  84.0 &  $65.2\pm1.2$    & $2.23\pm0.02$ \\
        W28            & 6.51    &  -0.29  &    $0.39\pm0.02$ &  1231.3 &                   92.1 & 0.014 & 0.013 &  30.4 &  $55.9\pm1.8$    & $2.65\pm0.03$ \\
        W30            & 8.61    &  -0.20  &    $0.36\pm0.04$ &   458.1 &                  66.0  & 0.020 & 0.018 &  14.1 &  $30.0\pm1.8$    & $2.58\pm0.06$ \\
        W44            & 34.68   &  -0.42  &    $0.34\pm0.02$ &  1387.6 &                  111.3 & 0.009 & 0.009 & -39.4 &  $74.7\pm1.0$    & $2.67\pm0.01$ \\
        W51C           & 49.12   &  -0.44  &    $0.28\pm0.02$ &  2174.7 &                   87.6 & 0.010 & 0.010 &  59.4 &  $41.6\pm1.3$    & $2.38\pm0.04$ \\
        \hline
        \multicolumn{11}{l}{$E > 10\gev$} \\
        \hline
        HESS J1825-137 &   17.58 &   -0.46 &    $0.64\pm0.08$ &    85.6 &                   62.5 & 0.050 & 0.045 & 33.6  &     $19.4\pm1.4$ & $2.27\pm0.07$ \\
        MSH 15-52      &  320.39 &   -1.22 &    $0.20\pm0.06$ &    73.7 &                    9.5 & 0.033 & 0.032 & 3.9   &      $4.7\pm0.7$ & $1.88\pm0.12$ \\
      \label{known_extended_sources}
      \end{tabular}
      \caption{The fit results for the known extended sources that
      were found by our analysis. The top results were found using an
      energy range between $1\gev$ and $100\gev$.  The lower analysis
      was done using events between $10\gev$ and $100\gev$.  The quoted
      flux is in units of $10^{-9}\ph/\cm^2/\s$ and is the flux in the
      fit energy range (either 1\gev to 100\gev or 10\gev to 100\gev).
      All sources were fit using a uniform disk spatial template. The
      localiztaion and extension were found using \pointlike while the TS
      and spectral values were calculated using \gtlike.  The TS value was
      computed using the assumed disk morphology. The Centaurus A Lobes,
      the Cygnus Loop, and Vela X were not found to be extended by this
      too. The reasons for this are discussed in~\ref{validate_known}.
      }
    \end{centering}
\end{sidewaystable}
